News Across Languages - Cross-Lingual Document Similarity and Event Tracking

### 2015-09-04 rupnik2015.tex

\textbf{Matrix factorization}. Non-negative matrix factorization-based \cite{nonnegfactor_lsi}, Cross-Lingual Latent Semantic Indexing CL-LSI\shortcite{cl_lsi}\cite{multilingualBook}, Canonical Correlation Analysis (CCA) \cite{Hotelling}, Oriented Principal Component Analysis \cite{platt2010translingual}. The quadratic time and space dependency of the OPCA method makes it impractical for large scale purposes. In addition, OPCA forces the vocabulary sizes for all languages to be the same, which is less intuitive. For our setting, the method in \cite{nonnegfactor_lsi} has a prohibitively high computational cost when building models (it uses dense matrices of dimensions that is the product of training set size and vocabulary size). Our proposed approach combines CCA and CL-LSI. Another closely related method is Explicit Semantic Analysis (ESA) \cite{ESA}, which uses Wikipedia (as do we in the current work) to compare documents. The authors of ESA compare it to CL-LSI and find that CL-LSI can outperform ESA in an information retrieval, but is costlier to optimize over a large corpus (ESA requires no training). We find that the scalability argument does not apply in our case: based on advances in numerical linear algebra we can solve large CL-LSI problems (millions of documents as oposed to 10,000 document limit reported in \cite{ESA}).

>>>>

\textbf{Matrix factorization}. Non-negative matrix factorization based \cite{nonnegfactor_lsi}, Cross-Lingual Latent Semantic Indexing CL-LSI\shortcite{cl_lsi}\cite{multilingualBook}, Canonical Correlation Analysis (CCA) \cite{Hotelling}, Oriented Principal Component Analysis \cite{platt2010translingual}. The quadratic time and space dependency of the OPCA method makes it impractical for large scale purposes. In addition, OPCA forces the vocabulary sizes for all languages to be the same, which is less intuitive. For our setting, the method in \cite{nonnegfactor_lsi} has a prohibitively high computational cost when building models (it uses dense matrices of dimensions that is the product of training set size and vocabulary size). Our proposed approach combines CCA and CL-LSI. Another closely related method is Explicit Semantic Analysis (ESA) \cite{ESA}, which uses Wikipedia (as do we in the current work) to compare documents. The authors of ESA compare it to CL-LSI and find that CL-LSI can outperform ESA in an information retrieval, but is costlier to optimize over a large corpus (ESA requires no training). We find that the scalability argument does not apply in our case: based on advances in numerical linear algebra we can solve large CL-LSI problems (millions of documents as opposed to 10,000 document limit reported in \cite{ESA}).
